---
title: "Lab 4A"
author: "Joe Minton"
date: "09/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
```

```{r}
surveys_complete <- read.csv("data/surveys_complete.csv")
```

## To build a ggplot, use the function format:
ggplot2(data = <DATA>, mapping = aes (<MAPPINGS>)) + <GEOM_FUNCTION> ()
```{r}
ggplot(data = surveys_complete)

## define aesthetic mapping (using aes function), by selecting variables to be potted and spcify how to present them (e.g., x/y poistions or characteristics or size, shape, color)

ggplot(data = surveys_complete, mapping = aes(x=weight, y=hindfoot_length))

## add "geoms" - graphical representations of the data in the plot (point, lines, bars). ggplot2 offers many different geoms

ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point()
```
```

## Assign plot to a variable
```{r}
surveys_plot <- ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length))
```

## Draw the plot
```{r}
surveys_plot + geom_point()
```

## Challenge 1: Allows hexagonal binning
```{r}
install.packages("hexbin")
library("hexbin")

surveys_plot <- ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length))

surveys_plot +
  geom_hex ()
  
## The strengths of the hexagonal has a key that shows a color code for regions of concentration (e.g., more point, the ligher blue it gets). For weaknesses, outliers are not as apparent and the shape slightly changes, showing a condensed graph. 
```

## To begin building plot, we start by defining the dataset, the axes, and choose a geom:
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point()
```

## Transparency (alpha) can be added to avoid overplotting
```{r}
ggplot(data = surveys_complete, aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1)
```

## To add colors for all points
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, color = "blue")
```

## To add colors to indivdual points, you cause use color and labeling that specific vector
```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) +
    geom_point(alpha = 0.1, aes(color = species_id))
```

## Challenge Set 2
```{r}
## Use what you learned to create a scatter plot of weight over species with the plot types showing in different colors. Is this a good way to show data?

ggplot(data = surveys_complete, 
       mapping = aes(x = species_id, y = weight)) +
   geom_point(aes(color = plot_type))
```

## Boxplots
```{r}
## Visualizing weight distribution within each species
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot()
```

## Adding points to the boxplot, we can see a better idea of number of measurements and of their distribution
```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_boxplot(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")
    
## Jitter adds the points on top of the box plot, showing the range, means, quartiles, etc.
```

## Challenge Set 3
```{r}
## violin plot or bean plot can show the shape of the distribution
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_violin(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato")

## changing the scale of the axes can show a better distribution of the data

ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) +
    geom_violin(alpha = 0) +
    geom_jitter(alpha = 0.3, color = "tomato") +
    scale_y_log10()

## making a new plot to explore the distribution of another variable within each species

ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
geom_jitter(aes(color = plot_id)) +
geom_boxplot(color = "pink")

## changing plot_id from integer to factor

ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) +
geom_jitter(aes(color = factor(c(plot_id)))) +
geom_boxplot(color = "black")
```