---
title: "Lab 9x: Programming in R-Control Structures and Functions"
name: "Joe Minton"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Control Structures
```{r}
# if and else: testing a condition and acting on it
# for: execute a loop a fixed number of times
# while: execute a loop while a condition is true
```

# Conditional statements with if-else
```{r, eval = FALSE}
# test a condition and act on it depending on whether it is true or false
if(<conditional statement>){
  ## do something
}
# Continue with rest of code
```

```{r, eval = FALSE}
# above code does nothing if condition is false; must have action to execute when condition is false, then you need an else clause
if(<conditional statement>){
  # do something
}
else{
  # do something else
}
```
```{r, eval = FALSE, warning=FALSE}
# can have series of tests by following the initial if with any number of else ifs

if(<conditional statement>) {
  # do something
} else if(<conditional statement>) {
  # do something different
} else {
  # do something else
}
```

# example of valid if/else structure
```{r, warning = FALSE, eval = FALSE}
# Generate a uniform random number
x <- runif(2,0,20)
if (x > 2) {
  y < - 20
} else {
  y <- 0
}
```

## Important Operators 
```{r}
# Comparison Operators
# == (equal)
# != (not equal)
# > (greater than)
# >= (greater than or equal)
# < (less than)
# <= (less than or equal)

# Logical Operators
# & (and)
# | (or)
# ! (not)

## for Loops
# in R, for loops take an interator variable and assign it successive values from a sequence or vector
```

```{r}
for(i in 0:14){
  print(i)
}
```
# loop takes i variable and in each iteration of the loop gives it values 0-14 and executes the code within brackets
```{r}
y <- c("j", "k", "l", "m")
for(i in 1:4)  {
  print(y[i])
}
```

```{r}
# do not need to use index variable
for(letter in y) {
  print(letter)
}
```

```{r}
# one line loops do not need curly braces
for(i in 1:4) print(y[i])
```

## Nested for loops
```{r}
# for loops can be nested inside of each other
```

```{r}
q <- matrix(1:10, 4, 5)
for(i in seq_len(nrow(q))) {
  for (j in seq_len(ncol(q))) {
    print(q[i, j])
  }
}
```

```{r}
# use nested loops for multidimensional data structures

## while loops
# while loops begin by testing a condition; if true, then execute loop body; loops exits when condition is false
```

```{r}
count <- 0
while(count < 14) {
  print(count)
  count <- count +2 
}
```
# while loops can result in infinitie loops if not wrtten properly

```{r}
z <- 10
set.seed(2)
while(z >= 6 && z <= 20) {
  coin <- rbinom(2, 2, 1)
  
  if(coin == 2){
    z <- z +2
  } else {
    z <- z - 2
  }
}
print(z)
```

## Functions
```{r}
# first class objects, which means that they can be treated much like any other R object

# first function
```

```{r}
# defined by using function()

 f <- function() {
   ## this is an empty function
 }
 ## functions have theur own class
class(f)
  ## execute function
 f()
```

# create function that has a non-trivial function body
```{r}
f <- function() {
  cat("Go, UMass!\n")
}
f()
```

# function arguements
```{r}
# can say how many time to run argument
f <- function(num) {
  for(i in seq_len(num)) {
    cat("Go, UMass!\n")
  }
}
f(3)
```
# tells you how many characters are in a function
```{r}
f <- function(num) {
  go <- "Go, UMass!\n"
  for(i in seq_len(num)) {
    cat(go)
  }
  chars <- nchar(go) * num
  chars
}
umassmotto <- f(3)
print(umassmotto)
```
# if num not specified, R provides an error
```{r, warning = FALSE, eval = FALSE}
f()
```
# can prevent error by setting default value to num
```{r}
f <- function(num = 2) {
  go <- "Go, UMass!\n"
  for(i in seq_len(num)) {
    cat(go)
  }
  chars <- nchar(go) * num
  chars
}
f()
f(4)
```

# functions have named arguments
```{r}
f(num = 2)
```









