---
title: "Lab 9: Programming in R-Regular Expression"
author: "Joe Minton"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regular Expression
```{r}
# a sequence of characters that define a search pattern; usually, such patterns are used by string-searching algorithms for "find" or "find and replace" operations or for input validation
```
# Base R Functions 
```{r}
# to deal with regular expressions:
# * grep(), grepl(): These functions search for matches of a regular expression/pattern in a character vector. grep() returns the indices into the character vector that contain a match or the specific strings that happen to have the match grepl() returns a TRUE/FALSE vector indicating which elements of the character vector contain a match
# * nregexpr(), gregexpr(): Search a character vector for regular expression matches and return the indices of the string where the match begins and the length of the match
# * sub(), gsub(): Search a character vector for regular expression matches and replace that match with another string
# * regexec(): This function searches a character vector for a regular expression, much like regexpr(), but it will additionally return the locations of any parenthesized sub-expressions. Probably easier to explain through demonstration.

# The stringr Package
#str_subset() is much like grep(value = TRUE) and returns a character vector of strings that contain a given match
# * str_detect() is essentially equivalent grepl()
# * str_extract() plays the role of regexpr() and regmatches(), extracting the matches from the output
# * str_match() does the job of regexec() by provide a matrix containing the parenthesized sub-expressions
```

##  14.3 Matching Patterns with Regular Expressions
```{r}
library(stringr)
library(tidyverse)
```
# 14.3.1 Basic matches
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")

# take character vector and regular expression and show you how they match
```

```{r}
str_view(x, ".a.")
# will search for an a with surrounding letters on both sides
```

```{r}
# \\ allows for searching for a character, such as ! ? . , 

# to create regular expression, need \\
dot <- "\\."
# expression itself on contains one
writeLines(dot)
# tels R to be explicit
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```

```{r}
# to find one \, you need four \ (\\\\)
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

# Exercsie 1 from 14.3.1
```{r}
# 1. "\" is an escape character in regular expression, finding a specific term; "\\" follwoed by a "." creates the regular expression; "\\\" creates a regular expression with the escape character without something to search for; "\\\\" creates a regular expression with the escape character "\", allowing it to search "\"
```

# 14.3.2 Anchors
```{r}
# use anchors to start or end a string being searched for:
# * ^ to match the start of the string
# * $ to match the end of the string
```

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```

```{r}
str_view(x, "a$")
```

```{r}
# to force a regular expression to only match a complete string, anchor both with ^ and $
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")

str_view(x, "^apple$")
```
# Exercise 1 from 14.3.2
```{r}
# to find "$^$", you can use "^\\$\\^\\$$"

# 14.3.3 Character Classes and Alternatives
# to find special patterns:
# * \d : matches any digit
# * \s : matches any whitespace (e.g. space, tab, newline)
# * [abc] : matches a, b, or c
# * [^abc] : matches anything except a, b, or c
# remember, to escape the \ for the string, use \\d or \\s
```

```{r}
# look for a literal character that normally has special meaning in a regular expression
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```

```{r}
str_view(x, ".[*]c")

str_view(x, "a[ ]")
```

```{r}
# This works for most (but not all) regex metacharacters: $ . | ? * + ( ) [ {. Unfortunately, a few characters have special meaning even inside a character class and must be handled with backslash escapes: ] \ ^ and -.
```

```{r}
# "|" can be used to pick beween one or more alternative patterns

x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

# Exercise 1 from 14.3.3
```{r}
str_subset(stringr::words, "^[aeiou]")
```

```{r}
str_view(stringr::words, "[aeiou]", match=FALSE)
```

```{r}
str_subset(c("ed", stringr::words), "(^|[^e])ed$")
```

```{r}
str_subset(stringr::words, "i(ng|se)$")
```
# 14.3.4 Repeition
```{r}
# controlling how many times a pattern matches:
# * ?: 0 or 1
# * +: 1 or more
# * : 0 or more
```

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```

```{r}
str_view(x, "CC+")
```

```{r}
str_view(x, 'C[LX]+')
```

```{r}
# can specify the number of matches precisely
# * {n}: exactly n
# * {n,}: n or more
# * {,m}: at most m
# * {n,m}: between n and m
```

```{r}
str_view(x, "C{2}")
```

```{r}
str_view(x, "C{2,}")
```

```{r}
# put a ? to match the shortest string possible
```

```{r}
str_view(x, "C{2,3}?")
```

```{r}
str_view(x, 'C[LX]+?')
```

# Exercise 1 from 14.3.4
```{r}
# Within {n,m} form, ? matches to the shortest string; + would allow for one additional character to be added; * would match the exact string
```

# 14.3.5 Grouping and backreferences
```{r}
# parentheses can create a numbered capturing group (number 1, 2, etc.): a capturing group stores the part of the string matched by the part of the regular expression inside the paratheses
# can use backreference, \1, \2, etc., to refer back to the same text previosuly matched by capturing group
```

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

# Exercise 1 from 14.3.5
```{r}
# (.)\1\1: matches to a character that shows up 3 times in a row (e.g., zzz)
# (.)(.)\\2\\1: matches a pair of characters followed by the same pair of characters in reversed order (e.g., cddc)
# (..)\1: matches any two repeated characters (e.g., v4v4)
# (.).\\1.\\1: matches a character by any character, the original character, another character, and the original again (e.g., xzxyx)
# (.)(.)(.).*\\3\\2\\: three matched characters follwoed by 0+ characters of any kind followed by the same three characters in reverse order (e.g., dfg9gfd)
```

## 14.4 Tools

# 14.4.1 Detect Matches
```{r}
# use str_detect() to see if a character vector matches a pattern
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

```{r}
# how many common words start with t?
sum(str_detect(words, "^t"))
```
```{r}
# What proportion of common words end with a vowel?
mean(str_detect(words, "[aeiou]$"))
```

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
```

```{r}
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

```{r}
# A common use of str_detect() is to select the elements that match a pattern. You can do this with logical subsetting, or the convenient str_subset() wrapper:
```

```{r}
words[str_detect(words, "x$")]
```

```{r}
str_subset(words, "x$")
```

```{r}
# want to filter 
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```

```{r}
# str_count(): rather than a simple yes or no, it tells you how many matches there are in a string
```

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```

```{r}
# On average, how many vowels per word?
mean(str_count(words, "[aeiou]"))
```

```{r}
# it is typical to see str_count() with mutate()
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

```{r}
str_count("abababa", "aba")
```

```{r}
str_view_all("abababa", "aba")
```

# Exercise 1 from 14.4.1

# * words that start/end with x
```{r}
words[str_detect(words, "^x|x$")]
start_with_x <- str_detect(words, "^x")
end_with_x <- str_detect(words, "x$")
words[start_with_x | end_with_x]
```

# * words starting with vowel and ending with consonant
```{r}
str_subset(words, "^[aeiou].*[^aeiou]$") %>% head()
start_with_vowel <- str_detect(words, "^[aeiou]")
end_with_consonant <- str_detect(words, "[^aeiou]$")
words[start_with_vowel & end_with_consonant] %>% head()
```

# * Are there any words that contain at least one of each different vowel? none.

# 14.4.2 Extract Matches
```{r}

# str_extract(): extracts actual text of a match
```

```{r}
length(sentences)
head(sentences)
```

```{r}
# to find sentences that contain a color, mke the expression
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

```{r}
# extracting colours
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)

has_colour
```

# str_extract() only extracts the first match, so to extract more:
```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```

```{r}
str_extract(more, colour_match)
```

```{r}
# use str_extract_all() to get all matches
str_extract_all(more, colour_match)
```

```{r}
# using smplify = TURE, str_extract_all() wil lreturn a matrix with short matches expanded to the same length as the longest
str_extract_all(more, colour_match, simplify = TRUE)

x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

# Exercise 1 from 14.4.2
```{r}
colour_matched <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")

more <- sentences[str_count(sentences, colour_matched) > 1]
str_view_all(more, colour_matched, match = TRUE)
```

# 14.4.3 Grouped matches
```{r}
# parantheses can be used for clarifying precedence and for backreferences when matching; also, they can be used to extract parts of a complex match
```

```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```

```{r}
# str_match() gives each individual column in a matrix, with columns for each group
has_noun %>% 
  str_match(noun)
```

```{r}
# can use tidyr::extract() if data is in a tibble; have to name the matches 
```

```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

```{r}
# if you want all matches for each string, use str_match_all
```

# Exercise 1 from 14.4.3
```{r}
numbs <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numbs)] %>%
  str_extract(numbs)
```

# 14.4.4 Replacing matches
```{r}
# str_replace() and str_replace_all() permit replacing matches with new strings
```

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```

```{r}
str_replace_all(x, "[aeiou]", "-")
```

```{r}
# with str_replacce_all(), you can perform mulitple replacements by supplying a name vector
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

```{r}
# can backreference to insert components of the match
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

# Exercise 1 from 14.4.4
```{r}
str_replace_all("first/second/third", "/", "\\\\")
```

# 14.4.5 Splitting
```{r}
#  use str_split to split a string up into pieces
```

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

```{r}
# can extract first element from length-1 vector
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

```{r}
# can use simplify = TRUE to return a matrix
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```


```{r}
# can set a maximum number of pieces 
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

# can split strings into characters, lines, sentence, and words using boundary()
```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]

str_split(x, boundary("word"))[[1]]
```

# Exercise 1 from 14.4.5
```{r}
x <- c("apples, pears, and bananas")
str_split(x, ", +(and +)?")[[1]]
```




